
```{r include=FALSE}
library(ggplot2)
library(rmarkdown)
library(tidyverse)
library(corrplot)
library(psych)
library(gridExtra)


data_dir<-"merged_SVI/"

GC_SVI_file<-"merged_CDC_GC.csv"

GC_SVI<-read.csv(paste0(data_dir,GC_SVI_file))
GC_SVI[GC_SVI==""]<-NA

# GC_SVI$Timestamp<-as.Date(GC_SVI$Timestamp, format ="%d-%b-%Y")

GC_SVI$Veteran[GC_SVI$Veteran=="Yes"]<-1
GC_SVI$Veteran[GC_SVI$Veteran!="1" | is.na(GC_SVI$Veteran)]<-0
GC_SVI$Veteran<-as.numeric(GC_SVI$Veteran)

GC_SVI$Incarcerated[GC_SVI$Incarcerated=="Yes"]<-1
GC_SVI$Incarcerated[GC_SVI$Incarcerated!="1"| is.na(GC_SVI$Incarcerated)]<-0
GC_SVI$Incarcerated<-as.numeric(GC_SVI$Incarcerated)

GC_SVI$Disability[GC_SVI$Disability=="Yes"]<-1
GC_SVI$Disability[GC_SVI$Disability!="1"| is.na(GC_SVI$Disability)]<-0
GC_SVI$Disability<-as.numeric(GC_SVI$Disability)

bed_req<-GC_SVI[,colnames(GC_SVI)[endsWith(colnames(GC_SVI), "BedReq")]]

for (i in 1:nrow(bed_req)){
  for (j in 1:ncol(bed_req)){
    if (is.na(bed_req[i,j])){
      bed_req[i,j]<-"0"
    }
    else{
    if (startsWith(bed_req[i,j], "Yes")){
      bed_req[i,j]<-"1"
    }
    else{
      bed_req[i,j]<-"0"
    }
    }
  }
}

##Assume that all NAs for those who have put down bed request are Nos

bed_req<-as.data.frame(lapply(bed_req,as.numeric))

GC_SVI<-cbind(GC_SVI, totalBedNeeded=rowSums(bed_req))
```

# Gathering some initial descriptive statistics on the numerical data
```{r echo=FALSE}

numer_subset<-colnames(GC_SVI)[sapply(GC_SVI, typeof)!="character" &
                                 !startsWith(colnames(GC_SVI), "More") &
                                 !startsWith(colnames(GC_SVI), "HH") &
                                 !startsWith(colnames(GC_SVI), "EPL") & 
                                 !startsWith(colnames(GC_SVI), "SPL") &
                                 !startsWith(colnames(GC_SVI), "RPL") &
                                 !startsWith(colnames(GC_SVI), "F") & 
                                 !startsWith(colnames(GC_SVI), "E_") & 
                                 !startsWith(colnames(GC_SVI), "M_") & 
                                 !startsWith(colnames(GC_SVI), "MP") ]

numer_subset<-numer_subset[-which(numer_subset %in% c("X.1", "X", "ClientZipCode",
                                                      "Timestamp"))]

numer_GCSVI<-GC_SVI[numer_subset]

describe(numer_GCSVI)
```

Some things worth noting about clients:

* Annual Income is right skewed; most annual incomes are small with a few large outliers
* NumAdultMale is slightly left skewed; most households have about one adult male but a lot of outliers with none
* Most beds needed are twin beds

Some things worth noting about SVI statistics of where clients live:

* Percent Poverty averages about 17%, but Percent Unemployment averages only about 5-6%
* About 10-12% have no high school diploma
* About 10-12% are single parents
* About 10-12% have a disability
* Average percent minority is over half
* About a fifth to a fourth are seventeen or younger
* On average, about 16% live in housing structures with more than 10 units per structure
* About 7-8% do not have a vehicle
* About 13% do not possess insurance

Some things of note for categorical variables:

* We see that a need for queen beds is positively correlated with those 
incarcerated and especially those with a disability
* We see some of the largest correlation values for the disability column
  + SVI measures of poverty, no high school diploma, speaking 
  English "less than well", living in apartment-like
  structures and crowded environments, having no vechicle, and
  not being insured

# Some correlation statistics:
Again some interesting things to note:

* NumChildren negatively correlated with SVI measures of poverty, living in apartment-like structures, and not having a vehicle
* Need for queen beds is slightly positively correlated with SVI measures of poverty and not having a vehicle
* Need for twin beds is slightly positively correlated with SVI measures of perecent of population aged less than or equal to 17, and percent of population living in mobile homes (in fact need for beds in general is correlated wth these SVI measures)

However, it is worth noting that SVI statistics may not be reliable for the given data.
For instance, the correlation between those with a disability among the clients
in the original dataset and the percent of population among that client with a disability
is quite small.

```{r echo=FALSE}

cor(numer_GCSVI, use="complete.obs")
```


```{r include=FALSE}

### Percent Median Income by Household Size
## Wake County
sixty_wake<-c(31920,36480,41040,45600,49260,52920)

sixty_wake_df<-data.frame(id=c(1:length(sixty_wake)),inc=sixty_wake)

##To get for household size 7,8 for Wake County, assuming a linear relationship
linear_sixty_wake<-lm(inc~id,data=sixty_wake_df)

sixty_wake_pred<-data.frame(id=c(7,8),
                            inc=predict(linear_sixty_wake, 
                                        newdata=data.frame(id=c(7,8))))

sixty_wake<-rbind(sixty_wake_df,sixty_wake_pred)$inc
 

#Raleigh
sixty_ral<-c(40200,45950,51700,57450,62050,66650,71200,75800)

eighty_ral<-c(53600,61250,68900,76550,82700,88800,94950,101050)


##Assuming a linear relationship between percent median income and percent level
## i.e. linear increase in annual income level for going from 30% to 60% median 
## income level

slope<-(eighty_ral-sixty_ral)/0.2

intercept<-sixty_ral-(slope*0.6)

thirty_ral<-(slope*0.3)+intercept

thirty_wake<-thirty_ral*mean(sixty_wake/sixty_ral)

##Adding 1s and 0s for in/out of % median income level group for 30% and 60% for
## both Wake County and Raleigh areas
ral_sixty<-c()
ral_thirty<-c()
wake_sixty<-c()
wake_thirty<-c()
for (i in 1:nrow(GC_SVI)){
  inc<-GC_SVI$AnnualIncomeAmount[i]
  numhouse<-GC_SVI$TotalHHNumber[i]
  if (!is.na(numhouse) &!is.na(inc)){
    if (0<numhouse & numhouse<9){
      if (inc<=sixty_ral[numhouse]){
        ral_sixty<-append(ral_sixty,1)
      }
      else{
        ral_sixty<-append(ral_sixty,0)
      }
      if (inc<=sixty_wake[numhouse]){
        wake_sixty<-append(wake_sixty,1)
      }
      else{
        wake_sixty<-append(wake_sixty,0)
      }
      if (inc<=thirty_ral[numhouse]){
        ral_thirty<-append(ral_thirty,1)
      }
      else{
        ral_thirty<-append(ral_thirty,0)
      }
      if (inc<=thirty_wake[numhouse]){
        wake_thirty<-append(wake_thirty,1)
      }
      else{
        wake_thirty<-append(wake_thirty,0)
      }
    }
    else{
      ral_sixty<-append(ral_sixty,NA)
      ral_thirty<-append(ral_thirty,NA)
      wake_sixty<-append(wake_sixty,NA)
      wake_thirty<-append(wake_thirty,NA)
    }
  }
  else{
    ral_sixty<-append(ral_sixty,NA)
    ral_thirty<-append(ral_thirty,NA)
    wake_sixty<-append(wake_sixty,NA)
    wake_thirty<-append(wake_thirty,NA)
  }
}

GCSVI_wInc<-cbind(GC_SVI,ral_sixty,ral_thirty,wake_sixty,wake_thirty)

GCSVI_wInc<-GCSVI_wInc[!is.na(GCSVI_wInc$ral_sixty),]

```



# Deeper Dive Into Correlated Relationships
```{r echo=FALSE}

ggplot(GCSVI_wInc, aes(x=NumChildren, y=EP_POV,color=as.character(NumChildren)))+geom_boxplot()+labs(x="Number of Children",y="Percent Poverty", title="Percent Poverty by Number of Children")+scale_color_discrete(name="Num Children", 
                       labels=c(0,1,2,3,4,5,6,7))
```
Some things to note:

* We see that that distribution among number of children is about the same for percent poverty
* Those with no children, however, often are among more impoverished areas, with a bit of a left-skew (indicating more distributed closer to more impoverished areas)

```{r echo=FALSE}
ggplot(GCSVI_wInc, aes(x=NumChildren, y=EP_MUNIT,color=as.character(NumChildren)))+geom_boxplot()+labs(x="Number of Children",y="Percent More than 10 Units", title="Percent More than 10 Units by Number of Children")+scale_color_discrete(name="Num Children", 
                       labels=c(0,1,2,3,4,5,6,7))
```
Some things to note:

* We see that the distribution among number of children varies; in general we see that having more children is associated with a smaller probability of living in a structure with more than ten housing units
  + This probability is pretty constant among after more than two kids

```{r echo=FALSE}
ggplot(GCSVI_wInc, aes(x=NumChildren, y=EP_NOVEH,color=as.character(NumChildren)))+geom_boxplot()+labs(x="Number of Children",y="Percent No Vehicle", title="Percent No Vehicle by Number of Children")+scale_color_discrete(name="Num Children", 
                       labels=c(0,1,2,3,4,5,6,7))
```
Some things to note:

* We see that that distribution among number of children is about the same, perhaps with a right skew (more towards a lower probability of having no vehicle) after the first child
  + There are a few clients that live in areas where the probability of having no vehicle is high no matter the number of children (long right tail)


```{r echo=FALSE}
ggplot(GCSVI_wInc, aes(x=QueenBeds, y=EP_POV,color=as.character(QueenBeds)))+geom_boxplot()+labs(x="Number Queen Beds Needed",y="Percent Poverty", title="Percent Poverty by Queen Beds Needed")+scale_color_discrete(name="Num Beds", 
                       labels=c(0,1,2,3,4,5,6,7))
```
Some things to note:

* We see that that distribution among number of queen beds varies. It is worth noting that:
  + Those that need one queen bed tend to live among more impoverished areas
  + It would be worth looking into how the need for a queen bed is distributed among those with a disability

```{r echo=FALSE}
ggplot(GCSVI_wInc, aes(x=QueenBeds, y=EP_NOVEH,color=as.character(QueenBeds)))+geom_boxplot()+labs(x="Number Queen Beds Needed",y="Percent No Vehicle", title="Percent No Vehicle by Queen Beds Needed")+scale_color_discrete(name="Num Beds", 
                       labels=c(0,1,2,3,4,5,6,7))
```
Some things to note:

* We see that that distribution among number of queen beds varies. It is worth noting that:
  + Those that need more than one queen bed tend to live in areas where the probability of having no vehicle tends to be lower (assuming that the need for three queen beds is an outlier)
  + Again, it would be worth looking into how the need for a queen bed is distributed among those with a disability

```{r echo=FALSE}
ggplot(GCSVI_wInc, aes(x=TwinBeds, y=EP_AGE17,color=as.character(TwinBeds)))+geom_boxplot()+labs(x="Number Twin Beds Needed",y="Percent Less Than 17", title="Percent Less Than 17 by Twin Beds Needed")+scale_color_discrete(name="Num Beds", 
                       labels=c(0,1,2,3,4,5,6,7))


```
Some things to note:

* We see that that distribution among number of twin beds varies. It is worth noting that overall, however, that the need for twin beds increases as one tends to live in areas with more people less than 17 years of age
