```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', fig.width = 12, fig.height = 8)
library(rgdal)
library(raster)
library(tidyverse)
library(readxl)
library(viridis)
library(rgeos)
library(sf)
library(osmdata)
library(ggmap)
library(here)
i_am("mapping_schools/mapping_schools.Rmd")
```

# Mapping Schools

```{r, echo = F}
### Load the cleaned data set
school_merge_csv <- read.csv(here("mapping_schools/school_merge.csv"))
### Calculate the proportion of students who are served by TGCP over the total number of students, in each school
count <- school_merge_csv %>% 
  group_by(SchoolName) %>%
  count()
Number.of.Students <- school_merge_csv %>% 
  distinct(SchoolName, .keep_all = TRUE)
df <- left_join(Number.of.Students, count)
df <- df %>% mutate(Proportion = n/Number.of.Students)
```

```{r, echo = F}
school_merge_df <- st_as_sf(df, coords = c("X", "Y"))
merge_shp <- st_set_crs(school_merge_df, 4326)
```

```{r, echo = F, eval = F}
### Load the street maps
mad_map <- get_map(getbb("Wake County North Carolina"), maptype = "toner-background")

big_streets <- getbb("Wake County North Carolina")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", "motorway_link", "primary_link")) %>%
  osmdata_sf()

med_streets <- getbb("Wake County North Carolina")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("secondary", "tertiary", "secondary_link", "tertiary_link")) %>%
  osmdata_sf()


small_streets <- getbb("Wake County North Carolina")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            "unclassified",
                            "service", "footway"
                  )) %>%
  osmdata_sf()

#river <- getbb("Wake County North Carolina")%>%
#  opq()%>%
#  add_osm_feature(key = "waterway", value = "river") %>%
#  osmdata_sf()

railway <- getbb("Wake County North Carolina")%>%
  opq()%>%
  add_osm_feature(key = "railway", value="rail") %>%
  osmdata_sf()

# the above objects take a long time to generate. I save them below so I can skip the re-computation.
save(mad_map, big_streets, med_streets, small_streets, railway, file = here("mapping_schools/map_features.RData"))
```

We show a plot of the schools (public, non-public) that students served by TGCP go to in Wake County, shown as red dots. Colleges and universities, daycare centers, and schools outside Wake County are not included in the plot. 

```{r, echo = F}
load(here("mapping_schools/map_features.RData"))

q <- ggmap(mad_map) +
  geom_sf(data = railway$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .2,
          linetype="dotdash",
          alpha = .5) +
  geom_sf(data = med_streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .3,
          alpha = .5) +
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "#666666",
          size = .2,
          alpha = .3) +
  geom_sf(data = big_streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .5)
```

```{r, echo = F}
q1 <- q + 
  theme_void() + 
  labs(title = "Map of Schools in Wake County") +
  geom_sf(data = merge_shp, inherit.aes = FALSE,
          color = "firebrick2") 
```

```{r, echo=F}
q1
```



A map of schools by proportion are also considered. For each school, the proportion of students who are served by TGCP over the total number of students are given below. 

```{r, echo = F}
q2 <- q + 
  theme_void() + 
  labs(title = "Map of Schools in Wake County, by Proportions") +
  geom_sf(data = merge_shp, inherit.aes = FALSE,
          aes(size = Proportion), alpha = 0.5,color = "firebrick2") 
```

```{r, echo = F}
q2
```