---
output:
  pdf_document: default
  html_document: default
---

<!-- ```{r, setup, include=FALSE} -->
<!-- options(knitr.graphics.auto_pdf = TRUE) -->
<!-- knitr::opts_chunk$set( -->
<!--   fig.height = 4, -->
<!--   fig.width = 6, -->
<!--   collapse = TRUE, -->
<!--   fig.align = "center", -->
<!--   dpi = 600 -->
<!-- ) -->
<!-- ``` -->

```{r warning=F, message=F, echo=F}
# Read data set and libraries
library(here)
i_am("tgcp-data/TGCP_mapping.rmd")
library(tidyverse)
library(choroplethr)
library(choroplethrZip)
library(choroplethrMaps)
library(usmap)
library(ggmap)
tgcp <- readRDS(here("cleaned_STATCOM_data_SVI.rds"))
tgcp <- tgcp[!is.na(tgcp$ClientZipCode),]
data(zip.regions)
north_carolina_zip <-
  zip.regions$region[zip.regions$state.name == "north carolina"]
# Define the zones which contain at least 5 clients
zip_code_summary <-
  tgcp[!is.na(tgcp$ClientZipCode),] %>% group_by(ClientZipCode) %>% summarise(num_clients = n())
zip_code_num_clients <-
  data.frame(region = as.character(zip_code_summary$ClientZipCode),
             value = zip_code_summary$num_clients)
atleast5_zip <-
  zip_code_num_clients$region[zip_code_num_clients$value > 4 &
                                !is.na(zip_code_num_clients$region)]
# Set reference Google map API
register_google("<your api key>")

# Define highlight function
highlight_county <- function(county_fips)
{
  data(county.map, package="choroplethrMaps", envir=environment())
  df = county.map[county.map$region %in% county_fips, ]
  geom_polygon(data=df, aes(long, lat, group = group), color = "Darkgreen", fill = NA, size = 1)
}
```

# Mapping TGCP variables

We focus on four variables from the TGCP client demographics data. The Wake county area is outlined in green. The mapping zone includes only the zip codes in which TGCP served at least five clients. For each variable, the proportion of clients served within the zip code with that particular characteristic is mapped. Clients with missing data are not included in the proportion.

<!-- ## Mapping population for the selected area -->

<!-- ```{r warning=F, message=F, echo=F} -->
<!-- data(df_pop_zip) -->
<!-- zip_choropleth(df_pop_zip, -->
<!--                zip_zoom = atleast5_zip, -->
<!--                title      = "Population estimates over zipcode", -->
<!--                legend     = "Population", -->
<!--                reference_map = F,   -->
<!--                num_colors = 9, -->
<!--              ) -->
<!-- ``` -->



## Mapping Disability
```{r warning=F, message=F, echo=F}
## Mapping Disability
zip_code_summary <-
  tgcp[tgcp$Disability != "N/A", ] %>% group_by(ClientZipCode) %>% # group_by just zip code
  summarise(num_Client = n(), num_Disability = sum(Disability == "Yes")) %>% # replace num_Disability by num_Client, add another summary for specific characteristic
  mutate(prop_Disability = num_Disability / num_Client) # get zip code specific proportion

# zip_code_num_disability <-
#   data.frame(
#     region = as.character(zip_code_summary$ClientZipCode),
#     value = zip_code_summary$num_Client
#   )

zip_code_prop__disability <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = round(zip_code_summary$prop_Disability, digits = 3)
  )

# zip_choropleth(
#   zip_code_num_disability,
#   zip_zoom = atleast5_zip,
#   title    = "Number of Disability",
#   legend   = "Number of Disability",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

zip_choropleth(
  zip_code_prop__disability,
  zip_zoom = atleast5_zip,
  title    = "Proportion of Disability",
  legend   = "Proportion of Disability",
  num_colors = 9,
  reference_map = F)+
  highlight_county(37183)
```


## Mapping Veteran
```{r warning=F, message=F, echo=F}
## Mapping Veteran
zip_code_summary <-
  tgcp[tgcp$Veteran != "N/A", ] %>% group_by(ClientZipCode) %>%
  summarise(num_Client = n(), num_Veteran = sum(Veteran == "Yes")) %>%
  mutate(prop_Veteran = num_Veteran / num_Client) 

# zip_code_num_Veteran <-
#   data.frame(
#     region = as.character(zip_code_summary$ClientZipCode),
#     value = zip_code_summary$num_Client
#   )

zip_code_prop__Veteran <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = round(zip_code_summary$prop_Veteran, digits = 3)
  )

# zip_choropleth(
#   zip_code_num_Veteran,
#   zip_zoom = atleast5_zip,
#   title    = "Number of Veteran",
#   legend   = "Number of Veteran",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

zip_choropleth(
  zip_code_prop__Veteran,
  zip_zoom = atleast5_zip,
  title    = "Proportion of Veteran",
  legend   = "Proportion of Veteran",
  num_colors = 9,
  reference_map = F)+
  highlight_county(37183)
```
<!-- ## Mapping COVID status -->
<!-- ```{r warning=F, message=F, echo=F} -->
<!-- ## Mapping COVID status -->
<!-- tgcp_COVID <- tgcp[!is.na(tgcp$COVID.19),] -->
<!-- zip_prep_summary <- -->
<!--   tgcp_COVID %>% group_by(ClientZipCode, COVID.19) %>% -->
<!--   summarise(num_COVID = n()) %>% -->
<!--   mutate(prop_COVID = num_COVID / sum(num_COVID)) -->
<!-- zip_code_summary <- zip_prep_summary %>% filter(COVID.19 == "COVID") -->
<!-- zip_code_num_COVID <- -->
<!--   data.frame(region = as.character(zip_code_summary$ClientZipCode), -->
<!--              value = zip_code_summary$num_COVID) -->

<!-- zip_code_prop__COVID <- -->
<!--   data.frame( -->
<!--     region = as.character(zip_code_summary$ClientZipCode), -->
<!--     value = round(zip_code_summary$prop_COVID, digits = 3) -->
<!--   ) -->

<!-- # zip_choropleth( -->
<!-- #   zip_code_num_COVID, -->
<!-- #   zip_zoom = atleast5_zip, -->
<!-- #   title    = "Number of COVID cases", -->
<!-- #   legend   = "Number of COVID cases", -->
<!-- #   num_colors = 9, -->
<!-- #   reference_map = F)+ -->
<!-- #   highlight_county(37183) -->

<!-- zip_choropleth( -->
<!--   zip_code_prop__COVID, -->
<!--   zip_zoom = atleast5_zip, -->
<!--   title    = "Proportion of COVID cases", -->
<!--   legend   = "Proportion of COVID cases", -->
<!--   num_colors = 9, -->
<!--   reference_map = F)+ -->
<!--   highlight_county(37183) -->
<!-- ``` -->



## Mapping Household Size

```{r warning=F, message=F, echo=F}
## Mapping TotalHHNumber
tgcp_TotalHHNumber <- tgcp[!is.na(tgcp$TotalHHNumber),]
tgcp_TotalHHNumber$TotalHHNumber <- as.numeric(tgcp_TotalHHNumber$TotalHHNumber)
zip_code_summary <-
  tgcp_TotalHHNumber %>% group_by(ClientZipCode) %>%
  summarise(num = n(), num_large = sum(TotalHHNumber >= 5), mean_hh = mean(TotalHHNumber)) %>%
  mutate(prop = num_large / num)

zip_code_num_TotalHHNumber <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = zip_code_summary$num_large
  )

zip_code_prop__TotalHHNumber <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = round(zip_code_summary$prop, digits = 3)
  )

zip_code_mean_TotalHHNumber <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = zip_code_summary$mean_hh
  )

# zip_choropleth(
#   zip_code_num_TotalHHNumber,
#   zip_zoom = atleast5_zip,
#   title    = "Number of large household",
#   legend   = "Number of large household",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

# zip_choropleth(
#   zip_code_prop__TotalHHNumber,
#   zip_zoom = atleast5_zip,
#   title    = "Proportion of large household",
#   legend   = "Proportion of large household",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

zip_choropleth(
  zip_code_mean_TotalHHNumber,
  zip_zoom = atleast5_zip,
  title    = "Mean Household Size",
  legend   = "Mean Household Size",
  num_colors = 9,
  reference_map = F)+
  highlight_county(37183)
```


## Mapping Home Income: Very low income status

Very low income status refers to being below 30% area median income. 

```{r warning=F, message=F, echo=F}
## Mapping Homeincome
tgcp_Homeincome <- tgcp[which(tgcp$Homeincome != "Unknown"),]
zip_code_summary <-
  tgcp_Homeincome %>% group_by(ClientZipCode) %>%
  summarise(num_Client = n(), num_VLIncome = sum(Homeincome == "Very Low Income")) %>%
  mutate(prop_Homeincome = num_VLIncome / num_Client) 
zip_code_num_Homeincome <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = zip_code_summary$num_VLIncome
  )

zip_code_prop__Homeincome <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = round(zip_code_summary$prop_Homeincome, digits = 3)
  )

# zip_choropleth(
#   zip_code_num_Homeincome,
#   zip_zoom = atleast5_zip,
#   title    = "Number of very low income home",
#   legend   = "Number of very low income home",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

zip_choropleth(
  zip_code_prop__Homeincome,
  zip_zoom = atleast5_zip,
  title    = "Proportion of very low income home",
  legend   = "Proportion of very low income home",
  num_colors = 9,
  reference_map = F) +
  highlight_county(37183)
```


## Mapping Home Income: Low income status

Low income status refers to being between 30% to 60% area median income. 

```{r warning=F, message=F, echo=F}
## Mapping Homeincome
tgcp_Homeincome <- tgcp[which(tgcp$Homeincome != "Unknown"),]
zip_code_summary <-
  tgcp_Homeincome %>% group_by(ClientZipCode) %>%
  summarise(num_Client = n(), num_VLIncome = sum(Homeincome == "Low Income")) %>%
  mutate(prop_Homeincome = num_VLIncome / num_Client) 
zip_code_num_Homeincome <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = zip_code_summary$num_VLIncome
  )

zip_code_prop__Homeincome <-
  data.frame(
    region = as.character(zip_code_summary$ClientZipCode),
    value = round(zip_code_summary$prop_Homeincome, digits = 3)
  )

# zip_choropleth(
#   zip_code_num_Homeincome,
#   zip_zoom = atleast5_zip,
#   title    = "Number of low income home",
#   legend   = "Number of low income home",
#   num_colors = 9,
#   reference_map = F)+
#   highlight_county(37183)

zip_choropleth(
  zip_code_prop__Homeincome,
  zip_zoom = atleast5_zip,
  title    = "Proportion of low income home",
  legend   = "Proportion of low income home",
  num_colors = 9,
  reference_map = F)+
  highlight_county(37183)
```
